---
title: "Data Analysis of Wolbachia Data using Frequency Wolbachia term"
author: "Nick Tierney & Nick Golding"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    toc: true
    toc-location: left
    code-fold: true
    code-tools: true
self-contained: true
fig-format: "png"
fig-dpi: 300
fig-asp: 1.618
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = TRUE)
## target knits qmds in their own session, so load libraries here.
source(here::here("packages.R"))
## if you want access to the functions you've written
lapply(list.files(here::here("./R"), full.names = TRUE), source)
```

```{r}
#| label: load-targets
#| include: false
withr::with_dir(here::here(), {
  tar_load(
    c(
      release_site_info,
      release_control_sites,
      cases_lookup,
      cases,
      incidence_lookup,
      incidence,
      wolbachia_2023,
      draws_fw,
      means_fw,
      wolbachia_2023_model_results_fw,
      parameter_summary_fw,
      perc_change_draws_fw_vec,
      dharma_fw
      )
    )
})
```

## The data

```{r}
#| label: show-data
wolbachia_2023
glimpse(wolbachia_2023)
summary(wolbachia_2023)
```

## Dengue incidence in control and study sites, with credible intervals

```{r}
#| label: wolbachia-add-cluster
wolbachia_2023_model_results_cluster_fw <- wolbachia_2023_model_results_fw %>% 
  mutate(cluster = substr(ministry_code, 1, 1),
         .after = ministry_code)
```

```{r}
#| label: show-panel-plot-1
#| fig-height: 50
#| fig-width: 8
wolbachia_2023_model_results_cluster_fw %>%
    filter(cluster == "C") %>%
    panel_plot()
```

```{r}
#| label: show-panel-plot-2
#| fig-height: 10
#| fig-width: 8
wolbachia_2023_model_results_cluster_fw %>%
    filter(cluster == "W") %>%
    panel_plot()
```

## MCMC Summary and Diagnostics

We have just done some preliminary modelling with 1000 draws - the model takes a little while to fit (about 1 hour), so before we commit to more samples we wanted to present the results so far.

### Trace plot and density

```{r}
#| label: plot-draws
#| fig-height: 10
#| fig-width: 8
plot(draws_fw)
```

::: {.panel-tabset}

### Draws Summary

```{r}
#| label: summary-chains
summary(draws_fw)
```

### Gelman Diagnostic

```{r}
#| label: diagnostics
coda::gelman.diag(draws_fw)
```

### Effective Sample Size

```{r}
#| label: diag-ess
coda::effectiveSize(draws_fw)
```

### Dharma Residuals Test

```{r}
#| label: dharma-test
#| fig-height: 3
#| fig-width: 9
testResiduals(dharma_fw)
```

:::

## Preliminary paper results

The posterior probability of a reduction due to intervention:

```{r}
#| label: post-prob-reduction
#| code-fold: false
# posterior probability of a reduction
## these are calculated as:
    # summarise parameter of interest
    # perc_change <- 100 * (exp(intervention_effect) - 1)
    # perc_change_draws <- calculate(perc_change, values = draws)
    # summary(perc_change_draws)
    # perc_change_draws_vec <- as.matrix(perc_change_draws)[, 1]

post_prob_reduction_fw <- mean(perc_change_draws_fw_vec < 0)
post_prob_reduction_fw
```

Quantiles and summary of the percentage change

```{r}
#| label: prob-quantiles
#| code-fold: false
# 95% CI
quantile(perc_change_draws_fw_vec, c(0.025, 0.975))
summary(perc_change_draws_fw_vec)
```

## Replication of "Figure 5" from 2019 paper

Dengue Reduction following Wolbachia Releases

```{r}
#| label: paper-plot-prep
intervention_label <- data.frame(
    ministry_code = "W01",
    date = as.Date("2020-09-01"),
    incidence_pred = 500
  )

# plot only intervention sites
wol_out_intervention_site_fw <- wolbachia_2023_model_results_fw %>%
    # subset to and rename release sites
    filter(str_detect(ministry_code, "^W")) %>%
    # add incidence data (observed and expected)
   mutate(incidence_pred = 100000 * post_pred_mean / population,
          incidence_lower = 100000 * post_pred_lower / population,
          incidence_upper = 100000 * post_pred_upper / population)
```


```{r}
last_dates_fw <- wol_out_intervention_site_fw %>% 
  select(ministry_code,
         date) %>% 
  group_by(ministry_code) %>% 
  filter(date == max(date)) %>% 
  ungroup()

intervention_rectangle_fw <- wol_out_intervention_site_fw %>% 
  select(ministry_code,
         start_release) %>% 
  distinct() %>% 
  left_join(last_dates_fw,
            by = "ministry_code") %>% 
  rename(start_date = start_release,
         finish_date = date)

```

```{r}
# get this working for one plot
wol_out_w1 <- wol_out_intervention_site_fw %>% 
  filter(ministry_code %in% c("W01")) %>% 
  group_by(ministry_code) %>% 
  mutate(fw_alt = fw * max(incidence),
         max_incidence = max(incidence),
         .after = fw) %>% 
  ungroup()

max(wol_out_w1$incidence)

ggplot(wol_out_w1,
       aes(x = date, 
           y = incidence_pred)) +
    geom_line(linewidth = 0.25) +
  geom_area(aes(y = fw_alt), fill = "lightblue", colour = "#83A4AE", alpha = 0.4) +
  geom_point(
      aes(x = date, y = incidence),
      col = grey(0.2),
      size = 0.25
    ) +
    geom_ribbon(aes(ymin = incidence_lower,
                    ymax = incidence_upper),
                linetype = 0,
                alpha = 0.2) +
    labs(
      x = "",
      y = "dengue incidence (per 100,000)"
    ) +
  scale_y_continuous(
    minor_breaks = NULL,
    sec.axis = sec_axis(~ . / max(wol_out_w1$incidence), name = "Wolbachia Frequency")
  ) +
    facet_wrap(~ministry_code, 
               ncol = 1,
               scales = "free") +
    geom_text(
      data = intervention_label,
      label = "post-release",
      size = 3.5
    ) +
    theme_minimal() +
    theme(strip.text = element_text(hjust = 0.5, size = 12))

```

```{r}
incidence_plot <- function(data){
  intervention_label <- data.frame(
    ministry_code = unique(data$ministry_code),
    date = as.Date("2020-09-01"),
    incidence_pred = 500
  )
  
  scale_factor <- max(data$incidence_upper) * 1.05
  
  plot_data <- data %>% 
    mutate(fw_alt = fw * max(incidence), .after = fw)
  
  ggp <- ggplot(plot_data,
       aes(x = date, 
           y = incidence_pred)) +
    geom_line(linewidth = 0.25) +
  geom_area(aes(y = fw_alt), 
            fill = "lightblue", 
            colour = "#83A4AE", 
            alpha = 0.4) +
  geom_point(
      aes(x = date, y = incidence),
      col = grey(0.2),
      size = 0.25
    ) +
    geom_ribbon(aes(ymin = incidence_lower,
                    ymax = incidence_upper),
                linetype = 0,
                alpha = 0.2) +
    labs(
      x = "",
      y = "dengue incidence (per 100,000)"
    ) +
  scale_y_continuous(
    minor_breaks = NULL,
    limits = c(0, scale_factor),
    breaks = scales::breaks_extended(n = 5),
    sec.axis = sec_axis(~ .x / scale_factor, name = "Wolb. Freq.",
                        breaks = c(0, 0.25, 0.5, 0.75,1)
                        )
  ) +
    facet_wrap(~ministry_code, 
               ncol = 1,
               scales = "free") +
    theme_minimal() +
    theme(strip.text = element_text(hjust = 0.5, size = 12))
  
  ggp
}
```

### plot as a list

```{r}
wol_out_intervention_site_fw %>%
  filter(ministry_code == "W01") %>%
  incidence_plot()

wol_out_intervention_site_fw %>%
  filter(ministry_code == "W02") %>%
  incidence_plot()

wol_out_intervention_site_fw %>%
  filter(ministry_code == "W03") %>%
  incidence_plot()

wol_out_intervention_site_fw %>%
  filter(ministry_code == "W03") %>%
  incidence_plot()
```


```{r}
wol_code_split <- split(wol_out_intervention_site_fw,
                        wol_out_intervention_site_fw$ministry_code)

incidence_plot_list <- map(wol_code_split, incidence_plot)
incidence_plot_list

# library(cowplot)
# 
# plot_grid(plotlist = incidence_plot_list, ncol = 1)
# pp <- plot_grid(p0, p1, ncol=1, rel_heights=c(0.1, 1))
```


### using patchwork

```{r}
#| label: paper-plot
#| fig-height: 40
#| fig-width: 6

library(patchwork)
patchwork::wrap_plots(
  incidence_plot_list,
  ncol = 1,
  guides = "collect"
  )
# gg_wol_out_intervention_site_fw <- ggplot(wol_out_intervention_site_fw,
#                                          aes(x = date, 
#                                              y = incidence_pred)) +
#   geom_rect(
#     data = intervention_rectangle_fw,
#     inherit.aes = FALSE,
#     aes(ymin = 0,
#         ymax = Inf,
#         xmin = start_date,
#         xmax = finish_date),
#       alpha = 0.4,
#       fill = "lightblue") +
#     geom_line(linewidth = 0.25) +
#     geom_ribbon(aes(ymin = incidence_lower,
#                     ymax = incidence_upper),
#                 linetype = 0,
#                 alpha = 0.2) +
#     geom_point(
#       aes(x = date, y = incidence),
#       col = grey(0.2),
#       size = 0.25
#     ) +
#     scale_y_continuous(minor_breaks = NULL) +
#     labs(
#       x = "",
#       y = "dengue incidence (per 100,000)"
#     ) +
#     facet_wrap(~ministry_code, 
#                ncol = 1,
#                scales = "free") +
#     geom_text(
#       data = intervention_label,
#       label = "post-release",
#       size = 3.5
#     ) +
#     theme_minimal() +
#     theme(strip.text = element_text(hjust = 0.5, size = 12))
# 
#   gg_wol_out_intervention_site_fw

```


## Reproducibility

<details><summary>Reproducibility receipt</summary>

```{r}
## datetime
Sys.time()

## repository
if(requireNamespace('git2r', quietly = TRUE)) {
  git2r::repository()
} else {
  c(
    system2("git", args = c("log", "--name-status", "-1"), stdout = TRUE),
    system2("git", args = c("remote", "-v"), stdout = TRUE)
  )
}

## session info
sessionInfo()
```

</details>
